### 2015-07-21 auspice/related.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

%A special case is that of structured names as in DNS.

%Due to the small update costs, 
%
%name to auspice \auspice\ mapping are expected to change at much smaller rates 
%
%. This is because name to \auspice\ provider mappings are expected to change infrequently

@@

	\textbf{Paxos overhead:}
}
\vspace{-0.1in}
\section{Related Work}
\label{sec:related}

@@

%\textbf{DNS enhancements:} 
%Since then, 
{\bf{DNS.}} Many have studied issues related to performance, scalability, load balancing, or denial-of-service vulnerabilities in DNS's resolution infrastructure \cite{Pappas,codons-paper,Brownlee,dnssec}. Several DHT-based alternatives have been put forward \cite{codons-paper,cox,DHTdns} and we compare against one representative proposal, Codons \cite{codons-paper}. In general, DHT-based designs are ideal for balancing load across servers, but are less well-suited to scenarios with a large number of service replicas that have to coordinate upon updates, and are at odds with scenarios requiring placement of replicas close to pocket of demand. In comparison, \auspice\ uses a planned placement approach.

@@

\textbf{Server selection.} Many prior systems have addressed the server selection problem with data or services replicated across a wide-area network. Examples include anycast services \cite{oasis,Bhattacharjee:1997:AA:839292.843045,donar} to map users to the best server based on server load or network path characteristics. These systems as well as CDNs and cloud hosting providers %\cite{AWS-load-balancing} 
share our goals of proximate server selection and load balance given a fixed placement of server replicas. \auspice\ differs in that it additionally considers replica placement itself as a degree of freedom in achieving latency or load balance.

\textbf{Dynamic placement.} We were unable to find prior systems that {\em automatically} reconfigure the {\em geo-distributed} replica locations of frequently {\em mutable objects} while preserving {\em consistency} (i.e., those satisfying all four italicized properties). However, reconfigurable placement has been studied for static or slow changing content \cite{gwertzman:95b} or within a single datacenter, or without replication. For example, Volley \cite{volley} optimizes the placement of mutable data objects based on the geo-distribution of accesses and is similar in spirit to \auspice\ in this respect, however it implicitly assumes a single replica for each object, so it does not have to worry about high update rates or replica coordination overhead. %As shown in  Section \ref{sec:eval}, creating multiple replicas of objects or services can significantly reduce user-perceived response times while also enhancing opportunities to balance load provided update cost is taken into account.
%\auspice\ is also similar in spirit to edge services hosted by CDNs. However, unlike \auspice, the geo-distributed locations of edge services as well as cloud-hosted services today are chosen manually and infrequently updated. %In comparison, \auspice\ automates geo-distributed placement of replicas, but the ``service'', a key-value store, is much simpler compared to black-box cloud-hosted services.

>>>>

%A special case is that of structured names as in DNS. 
%
%Due to the small update costs, 
%
%name to auspice \auspice\ mapping are expected to change at much smaller rates 
%
%. This is because name to \auspice\ provider mappings are expected to change infrequently

@@

	\textbf{Paxos overhead:}
}
\vspace{-0.1in}
\section{Related work}
\label{sec:related}

@@

%\textbf{DNS enhancements:} 
%Since then, 
{\bf{DNS:}} Many have studied issues related to performance, scalability, load balancing, or denial-of-service vulnerabilities in DNS's resolution infrastructure \cite{Pappas,codons-paper,Brownlee,dnssec}. Several DHT-based alternatives have been put forward \cite{codons-paper,cox,DHTdns} and we compare against one representative proposal, Codons \cite{codons-paper}. In general, DHT-based designs are ideal for balancing load across servers, but are less well-suited to scenarios with a large number of service replicas that have to coordinate upon updates, and are at odds with scenarios requiring placement of replicas close to pocket of demand. In comparison, \auspice\ uses a planned placement approach.

@@

\textbf{Request redirection:} Many prior systems have addressed the request redirection problem with data or services replicated across a wide-area network as discussed in the context of CDNs in Chapter \ref{ch:te-background}. Examples include anycast services \cite{oasis,Bhattacharjee:1997:AA:839292.843045,donar} to map users to the best server based on server load or network path characteristics. These systems as well as CDNs and cloud hosting providers share our goals of proximate request redirection and load balance given a fixed placement of server replicas. \auspice\ differs in that it additionally considers replica placement itself as a degree of freedom in achieving low latency or load balance.

\textbf{Dynamic placement:} We were unable to find prior systems that {\em automatically} reconfigure the {\em geo-distributed} replica locations of frequently {\em mutable objects} while preserving {\em consistency} (i.e., those satisfying all four italicized properties). However, reconfigurable placement has been studied for static or slow changing content \cite{gwertzman:95b} or within a single datacenter, or without replication. For example, Volley \cite{volley} optimizes the placement of mutable data objects based on the geo-distribution of accesses and is similar in spirit to \auspice\ in this respect, however it implicitly assumes a single replica for each object, so it does not have to worry about high update rates or replica coordination overhead. %As shown in  Section \ref{sec:eval}, creating multiple replicas of objects or services can significantly reduce user-perceived response times while also enhancing opportunities to balance load provided update cost is taken into account.
%\auspice\ is also similar in spirit to edge services hosted by CDNs. However, unlike \auspice, the geo-distributed locations of edge services as well as cloud-hosted services today are chosen manually and infrequently updated. %In comparison, \auspice\ automates geo-distributed placement of replicas, but the ``service'', a key-value store, is much simpler compared to black-box cloud-hosted services.

### 2015-07-21 background-new.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

The evaluation metric for ISP traffic engineering is a cost function that is dependent on the utilization of network links. A well-known cost function is maximum link utilization or MLU. A low link utilization is desirable to ISPs for two reasons. First, a low utilization of all links implies that the network is free from congestion hotspots. Second, it also implies that a network has more spare capacity to tolerate an increase in demand. Consider MLU as a capacity metric for example. If a traffic engineering scheme achieves an MLU of 0.25 for a given matrix, then it can tolerate up to a 4$\times$ surge in the load represented by the matrix.

@@

If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization. The routing thus computed is refered to as \emph{optimal traffic engineering} in literature. As it is impossible to have accurate knowledge of future traffic matrices, traffic engineering schemes either take a \emph{demand-aware} or a \emph{demand-oblivious} approach. A demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices. A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious  traffic engineering  is simpler to implement because it requires neither measuremnt nor periodic updating of network routing.

@@

\textbf{Request redirection:} In a CDN, request rediection occurs at two tiers - inter-datacenter and intra-datacenter. Our focus in this part of the thesis is on inter-datacenter request redirection because it influences wide-area traffic patterns that affect traffic engineering. We discuss intra datacenter redirection in the context of a Content Datcenter (Chapter \ref{ch:shrink}).

@@

Content delivery techniques can be classified into demand-aware and demand-oblivious similar to our classfication of traffic engineering schemes. We consider common content delivery techniques discussed above to be demand-oblivious since they do not require long term content-level measurement of demand but instead use simple online heuristics to make their decision.. In contrast, a demand-aware approach to  placement and redirection has also been studied, e.g., Applegate \emph{et al.} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

@@

\textbf{Interaction between traffic engineering and overlay routing:} Several results show the negative interaction between selfish overlay routing and network routing \cite{Roughgarden,selfishQiu}. Theoretical results indicate that the negative interaction could cause an arbitrary degradation in user perceived delay. Further studies using Internet-like traffic demands and topologies indicate that this interaction hurts traffic engineering metrics. While this body of work focuses on application adaptation to \emph{path diversity} created by overlay routing, our work focuses on \emph{location diversity} present in the Internet (Chapter \ref{ch:beyondmlu}). We do not consider overlay routing since our focus is primarily on intra-domain routing while overlay routing is most commonly used to mask inter-domain routing inefficiencies.

@@

A content-dominated traffic and the presence of content at multiple network locations changes the traditional ISP traffic enginering problem. In such a network, a traffic matrix is no longer an effective representation of demand. It can be changed via application-layer adaptation without any underlying change in demand as we demonstrate in this section. Demand is better expressed as a \emph{content matrix} in which each entry represents the demand for a content at a particular network location. A traffic matrix implicitly assumes a fixed source and a fixed destination for each traffic flow. But, a content matrix expresses the fact that the demand for a content at a particular destination can be served from any set of source locations. As we explain using examples below, the flexibility of serving the demand from any location raises questions on the usefulness of link-utilization based metrics that are commonly used in traffic engineering (Section \ref{sec:bg-poormlu}) as well on the importance of traffic engineering itself (Example \ref{sec:bg-ncdn-interaction}). These and other related questions motivate our work on traffic-engineering in content dominated networks.

@@

Link-utilization based metrics may not be a good predictor of user-perceived performance that depends on other factors such as propagtion delay, access link capacity and backbone link capacity also. Moreover, queuing-theoretic models show that only a high link utilization severely affects performance-critical metrics such as queuing delay and packet loss. For low to moderate link utilization that is common in today's ISPs, a reduction in link utilization may not yield a commensurate benefit in application peformance.

@@

Link-utilization based metrics such as MLU are poor metric for network capacity in the presence of location diversity. Unfortunately, as the example in Figure \ref{fig:3node-bg} shows, MLU is not a meaningful metric of capacity when application adaptation to location diversity determines the traffic matrix. 
Thus, location diversity necessitates a new capacity metric.

%Link-utilization based metrics may not be a good predictor of network capacity in scenarios where applications have the ability to download content from mulitple locations. We illustrate this point using an example. \tbd{complete the example, add figure}

@@

To appreciate how placement can shape traffic in an NCDN, consider the simple example in Figure~\ref{fig:NetworkExample}. Node $C$ has an object in its cache that is requested by end-users at nodes $A$ and $D$. Suppose that one unit of traffic needs to be routed from $C$ to $A$ and $0.5$ units  from $C$ to $D$ to satisfy the demand for that object. The routing that achieves the minimum MLU of $0.5$ to serve the demanded object is shown in the figure. Note that the routing that achieves the MLU of 0.5 is not possible with a simple, \unplanned\ scheme like hop-count as that would route all the traffic demand from $C$ to $A$ via $B$, resulting in an MLU of 1. Thus, a (\planned) traffic engineering scheme is necessary to achieve an MLU of 0.5.

@@

On the other hand, NCDNs can shape the traffic demand matrix by using a judicious placement and redirection scheme. Suppose that there is some space left in the content server's cache at node $B$ to accommodate an additional copy of the demanded object. By creating an additional copy of the object at $B$, the traffic demand of $A$ can be satisfied from $B$ and the demand of $D$ from $C$ achieving an MLU of $0.125$. In this case, judicious content placement decreased the MLU by a factor of $4$. Even more interestingly, this best MLU can be achieved using a simple routing scheme like hop-count routing while also improving user-perceived latency (assuming that the latency of link $BA$ is lower than that of the two-hop paths from $C$ to $A$). While this is a toy example, it shows that content placement flexibility reduces network cost and enables simpler routing. This interaction between placement and routing motivates us address the following questions for real content workloads and network topologies.

\emph{Research question: Does a joint optimization of placement, redirection and routing yield benefits over independently making these decisions using simple demand-oblivious schemes?}

>>>>

The evaluation metric for ISP traffic engineering is a cost function that is dependent on the utilization of network links. A well-known cost function is maximum link utilization or MLU \cite{rexford, COPE}. A low link utilization is desirable to ISPs for two reasons. First, a low utilization of all links implies that the network is free from congestion hotspots. Second, it also implies that a network has more spare capacity to tolerate an increase in demand. Consider MLU as a capacity metric for example. If a traffic engineering scheme achieves an MLU of 0.25 for a given matrix, then it can tolerate up to a 4$\times$ surge in the load represented by the matrix.

@@

If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization. The routing thus computed is refered to as \emph{optimal traffic engineering} in literature \cite{fortz2000internet,TEXCP}. As it is impossible to have accurate knowledge of future traffic matrices, traffic engineering schemes either take a \emph{demand-aware} or a \emph{demand-oblivious} approach. A demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices \cite{fortz2000internet,fortz2002traffic,COPE,MultiTM}. A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A common demand-oblivious technique is to use shortest path routing in which link weights are set to inverse of link capacities \cite{InvCap}, henceforth referred to as \invcap\ in this thesis. A demand-oblivious  traffic engineering is simple to implement because it does periodic traffic matrix measuremnt and routing configuration updates.

@@

\textbf{Request redirection:} In a CDN, request rediection occurs at two tiers - inter-datacenter and intra-datacenter. Our focus in this part of the thesis is on inter-datacenter request redirection because it influences wide-area traffic patterns that affect traffic engineering. We discuss intra datacenter redirection in the context of a content datcenter (Chapter \ref{ch:shrink}).

@@

Content delivery techniques can be classified into demand-aware and demand-oblivious similar to our classfication of traffic engineering schemes. We consider common content delivery techniques discussed above to be demand-oblivious since they do not require long term content-level measurement of demand but instead use simple online heuristics to make their decision.. In contrast, a demand-aware approach to  placement and redirection has also been studied, e.g., Applegate \emph{et al.} \cite{ATTVoD} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

@@

\textbf{Interaction between traffic engineering and overlay routing:} Several results show the negative interaction between selfish overlay routing and network routing \cite{Roughgarden,selfishQiu}. Theoretical results indicate that the negative interaction could cause an arbitrary degradation in user perceived delay. Further studies using Internet-like traffic demands and topologies indicate that this interaction hurts traffic engineering metrics. While this body of work focuses on application adaptation to \emph{path diversity} created by overlay routing, our work focuses on \emph{location diversity} present in the Internet (Chapter \ref{ch:beyondmlu}). We do not consider overlay routing since our focus is primarily on intra-domain routing while overlay routing is most commonly used to mask inter-domain routing inefficiencies \cite{Detour}.

@@

A content-dominated traffic and the presence of content at multiple network locations changes the traditional ISP traffic enginering problem. In such a network, a traffic matrix is no longer an effective representation of demand. It can be changed via application-layer adaptation without any underlying change in demand. Demand is better expressed as a \emph{content matrix} in which each entry represents the demand for a content at a particular network location. A traffic matrix implicitly assumes a fixed source and a fixed destination for each traffic flow. But, a content matrix expresses the fact that the demand for a content at a particular destination can be served from any set of source locations. As we explain using examples below, the flexibility of serving the demand from any location raises questions on the usefulness of link-utilization based metrics that are commonly used in traffic engineering (Section \ref{sec:bg-poormlu}) as well on the importance of traffic engineering itself (Example \ref{sec:bg-ncdn-interaction}). These and other related questions motivate our work on traffic-engineering in content dominated networks.

@@

Link-utilization based metrics may not be a good predictor of user-perceived performance that depends on other factors such as propagtion delay, access link capacity and backbone link capacity also. Moreover, queuing-theoretic models show that only a high link utilization severely affects performance-critical metrics such as queuing delay and packet loss \cite{mm1}. For low to moderate link utilization that is common in today's ISPs, a reduction in link utilization may not yield a commensurate benefit in application peformance.

@@

Link-utilization based metrics, in particular inverse of MLU, are a poor capacity metric in the presence of location diversity. In using inverse of MLU as a capacity, we implicitly assume that traffic matrices scale linearly as network traffic increases. 
But, this assumption may not hold if in a network where application adaptation to location diversity determines the traffic matrix as in the above example. Thus, location diversity necessitates a new capacity metric.

@@

To appreciate how placement can shape traffic in an NCDN, consider the simple example in Figure~\ref{fig:NetworkExample}. Node $C$ has an object in its cache that is requested by end-users at nodes $A$ and $D$. Suppose that one unit of traffic needs to be routed from $C$ to $A$ and $0.5$ units  from $C$ to $D$ to satisfy the demand for that object. The routing that achieves the minimum MLU of $0.5$ to serve the demanded object is shown in the figure. Note that the routing that achieves the MLU of 0.5 is not possible with a simple, \unplanned\ scheme like \invcap\ as that would route all the traffic demand from $C$ to $A$ via $B$, resulting in an MLU of 1. Thus, a (\planned) traffic engineering scheme is necessary to achieve an MLU of 0.5.

@@

On the other hand, NCDNs can shape the traffic demand matrix by using a judicious placement and redirection scheme. Suppose that there is some space left in the content server's cache at node $B$ to accommodate an additional copy of the demanded object. By creating an additional copy of the object at $B$, the traffic demand of $A$ can be satisfied from $B$ and the demand of $D$ from $C$ achieving an MLU of $0.125$. In this case, judicious content placement decreased the MLU by a factor of $4$. Even more interestingly, this best MLU can be achieved using a simple routing scheme like \invcap\ routing while also improving user-perceived latency (assuming that the latency of link $BA$ is lower than that of the two-hop paths from $C$ to $A$). While this is a toy example, it shows that content placement flexibility reduces network cost and enables simpler routing. This interaction between placement and routing motivates us address the following questions for real content workloads and network topologies.

\emph{Research question: Does a joint optimization of placement, redirection and routing yield benefits over independently making these decisions using simple demand-oblivious schemes in an NCDN?}

### 2015-07-21 conclusion.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

@@

We also studied the relative importance of optimizing placement, redirection and routing in this thesis. In an ISP network with content location diversity and a Network CDN, we showed that content placement flexibility signficantly reduces the value of a carefully engineered routing. In a global name service, a better redirection scheme does improve latencies, e.g., random-k outperforms DHT due to a better redirection scheme. However, a demand-aware placement can significantly improve cost and performance even on top of an effective redirection scheme, e.g.,  Auspice does signficantly better than random-k, which uses the same redirection scheme as Auspice, as well as managed DNS, which uses global anycast routing. Overall, these findings strengthen our thesis that placement is of key importance in infrastructure service design.
q

>>>>

\eat{\chapter{Future work}
The goal of this thesis has been to design infrastructure services for supporting a content-dominated, highly mobile Internet.

An overarching goal of our  future research is to design a cost-effective and high-performance infrastructure for delivering content and services that leverages resources at end users, network edge and cloud datacenters. Modern day Internet falls short of this goal in several aspects. First, the potential cost and performance benefits of cloud computing platforms are not always realized by today's applications. Second, the Internet has a limited deployment of in-network caches that are freely available to applications, and for this reason, expensive CDN services are needed to cache content near end users. Third, content delivery to mobile end hosts, an increasingly important group of users, is unsatisfactory due to various reasons including resource constraints on mobile devices and unpredictable cellular network performance. Below, We outline research topics towards addressing these shortcomings and realizing our  research goal.

\textbf{Network edge architecture:} Recent research (including our  work on NCDNs) has shown that much of the benefits due to pervasive in-network caching can be achieved by caching content only on the network edge. Still, many questions remain unanswered regarding the design of edge caches. What should be the granularity of cache placement – individual homes or a neighborhood, or a city? Which algorithms for content search are effective in locating content in nearby caches? What would be the computational and storage requirements for these caches? What, if any, changes do we need to the design of switches to interface with cache servers? Addressing these questions presents avenues for novel measurement studies, prototype development and deployment studies in organizations.

\textbf{Datacenter energy efficiency:} Energy is an important factor in the operational costs of a datacenter; therefore, energy efficiency is an important design goal for future networked systems. Recent research (including our  work on content datacenters) has shown that software-level solutions based on consolidation reduce datacenter energy use. But, many popular networked systems running inside datacenters were not designed to be energy efficient originally and hence might not interoperate well with consolidation-based energy optimization. A key open question is to evaluate the performance impact of energy optimization on various classes of networked systems such as network file systems, distributed data stores and distributed memory caches, and if needed, to re-architect these systems keeping energy efficiency as a primary design goal. We believe that this line of research would yield design principles that could be broadly applied to the design of energy-efficient networked systems.

\textbf{Mobile content delivery:} There is a pressing need to improve content delivery to mobile end hosts given the popularity of devices such as smartphones and tablets. We believe that a two-pronged approach is needed to do so. First approach is pushing content closer to end hosts by means of edge caches. The potential of edge caching inside cellular networks has not been exploited yet, and recent studies point to its benefits. Designing content delivery infrastructure inside cellular networks is an important research challenge that can potentially reduce infrastructure costs of cellular networks besides improving content delivery. Second approach is moving parts of computation from the end hosts and performing it on network edge or in cloud datacenters. But, evaluating the cost-benefit tradeoff of moving computation at an alternate location, and designing APIs that allow developers to do so are challenges that remain to be overcome.
}

@@

We also studied the relative importance of optimizing placement, redirection and routing in this thesis. In an ISP network with content location diversity and a Network CDN, we showed that content placement flexibility signficantly reduces the value of a carefully engineered routing. In a global name service, a better redirection scheme does improve latencies, e.g., random-k outperforms DHT due to a better redirection scheme. However, a demand-aware placement can significantly improve cost and performance even on top of an effective redirection scheme, e.g.,  Auspice does signficantly better than random-k, which uses the same redirection scheme as Auspice, as well as managed DNS, which uses global anycast routing. Overall, these findings strengthen our thesis that placement is of key importance in infrastructure service design.

### 2015-07-21 intro-new.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

Most of the Internet traffic is content, and most of the Internet connected hosts are mobile. A key example of the content-dominated nature of the Internet is that online video alone accounts for nearly 67\% of traffic today and is expected to grow to nearly 80\% by 2018 \cite{cisco-videogrowth}. The evidence of a highly mobile Internet is that close to 5 billion mobile devices connect to the Intenet today and they generate more traffic than wired devices \cite{cisco-vni}.

%It is estimated that a vast majority of network traffic on the Internet is content, with videos alone expected to account for 80\% of all traffic by 2018 . By 2020, the number of Internet connected mobile end hosts is expected to grow to nearly 10 billion and the total traffic originated by mobiles is poised to approach that by wired devices .

A content-dominated, highly mobile Internet needs several forms of infrastructure support. It needs network switches and links with sufficient capacity to carry traffic to end users. It needs servers to sustain and accelerate delivery of content. It also needs  infrastructure to help establish and maintain connections in the presence of high network mobility (or changing of addresses) of connection end-points. The focus of this thesis is to design \emph{services} that manage the resources on such infrastructures towards achieving their desired goals.

%the design of services that manage the infrastructure resources achieve the desired goals.
%This thesis focuses on the design of services that manage the infrastructure resources achieve the desired goals.

@@

\section{Infrastructure characteristics}
%	Characteristics of infrastructure}

@@

\textbf{Cost-intensive} There is a substantial captial and operational cost in running a large-scale infrastructure \cite{greenberg2008cost}, and which causes a sigificant reduction in the profit margins of the infrastructure owner. A prominent example of cost-intensive, low-profit infrastructures are Internet service provider (ISP) networks \cite{isp-low-profit}. Thus, a main goal of service design is to reduce cost while meeting performance and other constraints.

@@

\textbf{Locality-exhibiting workload:} Real-world workloads served by these infrastructures tend to exhibit significant geographic and temporal locality \cite{NCDN, youtubeUGC, vodP2Pbenefit, cellularvideotraffic}. Exploiting locality of demand to reduce infrastructure cost as well as user-perceived latencies is a key focus in our service designs.

@@

\begin{figure}

	\centering
	\includegraphics[scale=0.5]{fig/placement-vs-redirection.png}
	\caption{Placement vs. redirection: Content placement creates options for the redirection scheme to choose a nearby location for reducing user-perceived latencies.}	
	\label{fig:placement-redirection}
\end{figure}

\textbf{Placement vs. redirection:} Suppose a user requests a content that is placed only at a remote location as in Figure \ref{fig:placement-redirection} (left). Irrespective of the redirection scheme, a user will observe high latencies to fetch content from the remote location. Placement is more powerful that redirection becasue it can create more options for the redirection scheme, enabling it to choose a location that is nearby as shown in Figure \ref{fig:placement-redirection} (right). Thus, effective placement is a pre-requisite for a redirection scheme to provide low latencies.

@@

\subsection{Traffic engineering in content-dominated ISP networks}
\label{sec:intro-te}
In an ISP network with content-dominated traffic, traffic engineering decisions that compute network-layer routing are not isolated from traffic adaptation occuring at application-layer. For example, content placement and request redirection determine the traffic matrix and hence influence traffic engineering. While the interaction between network routing and application adaptation has been studied previously \cite{Roughgarden,selfishQiu,Jiang2009,JohariGameTheory, CATE, P4P}, the distinguishing aspect of our work it to study the role of placement strategies on this interaction.

@@

In today's content-dominated Internet, it is not surprising that datacenters and Points-of-Presence (PoPs) are dedicated to storing and serving content to end users. We call these datacenters and PoPs {\em content datacenters} (or, \cdc s). The three degrees of freedom -- placement, redirection and routing -- discussed in a wide-area setting previously, are also available inside a \cdc: content placement to select which servers to store on content on, request redirection to select which server inside a \cdc\ to send a request to, and network routing on \cdc\ topology.

Energy use is a key concern for \cdc\ operator, e.g., Akamai CDN, managing a large global network of \cdc's, which can comprise of 100K servers or even more. A potential approach to reduce energy is \emph{consolidation} in a \cdc. Consolidation of servers, which uses only a fraction \cdc\ servers for placement and redirection and turns remaining servers off, can reduce server energy use. Consolidation of network, which routes network traffic via only a fraction of switches and links and turns remaining switches and links off, can reduce network energy use. A key concern for operators seeking to deploy these techniques is to understand the tradeoff between energy use and performance, i.e., do energy saving benefits outweigh the performance penalty incurred?

Our work presents the first quantitative analysis of energy-performance tradeoff for a \cdc\ based on a real \cdc\ workload and present the design and implementation of a system, \shrink, to leverage this tradeoff. A key insight, supported via experiments, is that cache hit rates, despite server consolidation, remain close to an unconsolidated datacenter. A small reduction in hit rate helps mitigate the impact of consolidation on user-perceived latency. This finding is explained by the skewed content popularity obseved in real workloads, due to which working set size of content remains small compared to the storage available in a CDC. Our quantitative analysis shows that \shrink\ reduces energy use by 35\% over a baseline scheme that keeps entire datacenter always on while increasing the mean, 95-th \%-ile and 99-th \%-ile latencies by 8\%, 3\% and 15\% respectively.

While previous work has studied server and network consolidation in isolation, we show that a coordinated approach for server and network consolidation reduces \emph{network} energy use more than otherwise possible. We consider a simple \emph{network-aware server consolidation} scheme that selects the active servers in a left-to-right order in a topology. The same network consolidation scheme results in up to 42\% less network energy use when used with our network-aware server consolidation instead of a network-unaware server consolidation scheme that selects servers randomly.

@@

In the first part, which includes Chapter \ref{ch:te-background}, Chapter \ref{ch:beyondmlu} and Chapter \ref{ch:ncdn}, we present our research on traffic engineering in content-dominated ISP networks. Chapter \ref{ch:te-background} presents background on traffic engineering, content delivery and the interaction between traffic engineering and application-layer adaptation, and motivates the key research questions we address on this topic. Chapter \ref{ch:beyondmlu} presents a comparison of traffic engineering schemes in a network with location diversity of content. Chapter \ref{ch:ncdn} designs and evaluates traffic engineering and content delivery schemes in a Network CDN.

>>>>

Most of the Internet traffic is content, and most of the Internet connected hosts are mobile. A key example of the content-dominated nature of the Internet is that online video alone accounts for nearly 67\% of traffic today and is expected to grow to nearly 80\% by 2018 \cite{cisco-videogrowth}. The evidence of a highly mobile Internet is that close to 5 billion mobile devices connect to the Intenet today and they generate more traffic than wired devices \cite{cisco-vni}. Due to the prevalence of content traffic and mobile users, content-based services and mobile applications are cornerstones of today's Internet-based economy.

A content-dominated, highly mobile Internet needs several forms of infrastructure support to sustain itself. It needs network switches and links with sufficient capacity to carry traffic to end users. It needs servers to sustain and accelerate delivery of content. It also needs  infrastructure to help establish and maintain connections in the presence of high network mobility (or changing of addresses) of connection end-points. This thesis seeks to design \emph{infrastructure services} that manage the resources on such infrastructures towards achieving their desired goals.

%The design of services that manage the infrastructure resources achieve the desired goals. This thesis focuses on the design of services that manage the infrastructure resources achieve the desired goals.

@@

\section{Infrastructure characteristics}

@@

\textbf{Cost-intensive:} There is a substantial captial and operational cost in running a large-scale infrastructure \cite{greenberg2008cost}, and which causes a sigificant reduction in the profit margins of the infrastructure owner. A prominent example of cost-intensive, low-profit infrastructures are Internet service provider (ISP) networks \cite{isp-low-profit}. Thus, a main goal of service design is to reduce cost while meeting performance and other constraints.

@@

\textbf{Locality-exhibiting workload:} Real-world workloads served by these infrastructures tend to exhibit significant geographic and temporal locality \cite{NCDN, youtubeUGC, vodP2Pbenefit, cellularvideotraffic}. Exploiting locality of demand to reduce infrastructure cost as well as user-perceived latencies is another key focus in our service designs.

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{fig/placement-vs-redirection.png}
	\caption{Placement vs. redirection: Content placement creates options for the redirection scheme to choose a nearby location for reducing user-perceived latencies.}	
	\label{fig:placement-redirection}
\end{figure}

\textbf{Placement vs. redirection:} Suppose a user requests a content that is placed only at a remote location as in Figure \ref{fig:placement-redirection} (left). Irrespective of the redirection scheme, a user will observe high latencies to fetch content from the remote location. Placement is more powerful that redirection becasue it can create more options for the redirection scheme, enabling it to choose a location that is nearby as shown in Figure \ref{fig:placement-redirection} (right). Thus, effective placement is a prerequisite for a redirection scheme to provide low latencies.

@@

\subsection{Traffic engineering in content-dominated ISP networks}
\label{sec:intro-te}
In an ISP network with content-dominated traffic, traffic engineering decisions that compute the network layer routing are not isolated from traffic adaptation occuring at the application layer. For example, content placement and request redirection determine the traffic matrix and hence influence traffic engineering. While the interaction between network routing and application adaptation has been studied previously \cite{Roughgarden,selfishQiu,Jiang2009,JohariGameTheory, CATE, P4P}, a distinguishing aspect of our work it to study the role of placement schemes on this interaction.

@@

In today's content-dominated Internet, it is not surprising that many datacenters and Points-of-Presence (PoPs) are dedicated to storing and serving content to end users. We call these datacenters and PoPs {\em content datacenters} (or, \cdc s). The three degrees of freedom -- placement, redirection and routing -- discussed in a wide-area setting previously, are also available inside a \cdc: content placement to select which servers to store on content on, request redirection to select which server inside a \cdc\ to send a request to, and network routing on \cdc\ topology.

Energy use is a key concern for \cdc\ operator managing a large global network of \cdc's, which can comprise of 100K servers or even more \cite{akamai-facts,qureshi2009cutting}. A potential approach to reduce energy use is \emph{consolidation} in a \cdc. Consolidation of servers can reduce server energy use by using only a fraction of \cdc's servers for placement and redirection and turns remaining servers off \cite{mathew12}. Similarly, consolidation of network can reduce network energy use by routing network traffic via only a fraction of switches and links and turns remaining switches and links off \cite{response}. A key concern for operators seeking to deploy these techniques is to understand the tradeoff between energy use and performance, i.e., do energy saving benefits outweigh the user-perceived performance penalty incurred?

Our work presents the first quantitative analysis of the energy-performance tradeoff for a \cdc\ based on a real \cdc\ workload and present the design and implementation of a system, \shrink, to leverage this tradeoff. A key insight, supported via experiments, is that cache hit rates, despite server consolidation, remain close to an unconsolidated datacenter. A small reduction in hit rate helps mitigate the impact of consolidation on user-perceived latency. This finding is explained by the skewed content popularity obseved in real workloads, due to which working set size of content remains small compared to the storage available in a CDC. Our quantitative analysis shows that \shrink\ reduces energy use by 35\% over a baseline scheme that keeps entire datacenter always on while increasing the mean, 95-th \%-ile and 99-th \%-ile latencies by 8\%, 3\% and 15\% respectively.

Further, we show that a coordinated approach for server and network consolidation reduces \emph{network} energy use more than otherwise possible. In comparison, previous work has studied server and network consolidation in isolation. We consider a simple \emph{network-aware server consolidation} scheme that selects the active servers in a left-to-right order in a topology. The same network consolidation scheme results in up to 42\% less network energy use when used with our network-aware server consolidation instead of a network-unaware server consolidation scheme that selects servers randomly.

@@

The first part, which includes Chapter \ref{ch:te-background}, Chapter \ref{ch:beyondmlu} and Chapter \ref{ch:ncdn}, presents our research on traffic engineering in content-dominated ISP networks. Chapter \ref{ch:te-background} presents background on traffic engineering, content delivery and the interaction between traffic engineering and application-layer adaptation, and motivates the key research questions we address on this topic. Chapter \ref{ch:beyondmlu} presents a comparison of traffic engineering schemes in a network with location diversity of content. Chapter \ref{ch:ncdn} designs and evaluates traffic engineering and content delivery schemes in a Network CDN.

### 2015-07-21 shrink/analysis.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

We present a simple analytical model in order to quantify the relationship between energy savings and response-time inflation in \cdc s, and to gain a deeper understanding of the factors that determine this trade-off. The primary purpose of this model is expository, so we make a number of simplifying assumptions for ease of exposition first, and we progressively relax them in this and subsequent sections.

>>>>

We present a simple analytical model in order to quantify the relationship between energy savings and user-perceived latency inflation in \cdc s, and to gain a deeper understanding of the factors that determine this trade-off. The primary purpose of this model is expository, so we make a number of simplifying assumptions for ease of exposition first, and we progressively relax them in this and subsequent sections.

### 2015-07-21 shrink/intro.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

(1) \emph{Comparison to baseline:}  \shrink\ reduces energy use by 35\% over a baseline scheme that provisions resources according to the peak demand while increasing mean, 95-th \%-ile and 99-th \%-ile latency by 8\%, 3\% and 15\% respectively.

@@

(3) \emph{Comparison of network energy use:} When one-fourth of the servers are active, \shrink's network energy use is lower than a network-unaware server consolidation scheme by 37\% and 42\% on FatTree \cite{fattree} and VL2 \cite{vl2} respectively.

>>>>

(1) \emph{Comparison to baseline:}  \shrink\ reduces energy use by 35\% over a baseline scheme that provisions resources according to the peak demand while increasing the mean, 95-th \%-ile and 99-th \%-ile latencies by 8\%, 3\% and 15\% respectively.

@@

(3) \emph{Comparison of network energy use:} When one-fourth of the servers are active, \shrink's network energy use is lower than a network-unaware server consolidation scheme by 37\% and 42\% on the FatTree \cite{fattree} and the VL2 \cite{vl2} topology respectively.
